---
title: "Enhanced `deseason()`-ing with **Rcpp**"
author: "Florian Detsch"
date: "26.05.2015"
output: html_document
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 120)
```

The [`deseason`](https://github.com/environmentalinformatics-marburg/remote/blob/master/R/deseason.R) function included in the [R **remote** package](https://github.com/environmentalinformatics-marburg/remote) has recently been [reimplemented in **Rcpp**](https://github.com/environmentalinformatics-marburg/remote/blob/develop/R/deseasonC.R), resulting in a performance speed more than 15 times faster as compared to the initial, solely R-based version.   

### Speed test

#### Prerequisites

```{r, message=FALSE}
## required packages
# devtools::install_github("environmentalinformatics-marburg/remote", 
#                          ref = "develop")
library(remote)
library(microbenchmark)

# devtools::install_github("hadley/lineprof")
library(lineprof)

## data
data("australiaGPCP")
```

#### Base-R deseasoning

```{r dsn_r}
microbenchmark(
  deseason(australiaGPCP)
, times = 20L)
```

#### **Rcpp**-based deseasoning

```{r dsn_cpp}
microbenchmark(
  deseasonC(australiaGPCP)
, times = 20L)
```


### Line profiling

While the initial approach using `lapply` along with `raster::calc` generated a considerable overhead, calculation efforts could now be minimized by the use of a) base-C++ `for` loops and b) a set of custom *.cpp functions analogous to base-R `seq` and vector indexing.

#### Base-R deseasoning

```{r lineprof_r, message=FALSE}
lineprof(deseason(x = australiaGPCP))
```

#### **Rcpp**-based deseasoning

```{r lineprof_cpp, message=FALSE}
lineprof(deseasonC(x = australiaGPCP))
```